Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-11-10 draft.tex
revise section 4

If Restriction (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1. If Restrictions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of Restriction 3. Note here from Restriction (2) that sanctions can induce investment behavior by the patron that was ruled out under the Restrictions of Definition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.

Thus, even if we do not consider the sanctions to impose any direct costs on the home state (though in practice they likely do), it remains ambiguous whether sanctions will benefit the home state. Recall that in the status quo equilibrium, the home state's expected returns from war are lower than from a continuation of the status quo. If sanctions induce the secessionists to play Fight, this is a worse outcome for the home state than if the status quo had been allowed to persist.

@@

\begin{proposition}
Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities. The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.

@@

The proof of Proposition 3 is immediate. Although under the conditions of Proposition 1 the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting. The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.

Thus, propositions 2 and 3 imply that sanctions are both wealth destroying and violence increasing. The sanctions destroy wealth directly by damaging the economy of the secessionist region and lowering the secessionists' payoffs from the status quo. If the degradation of status quo payoffs are not offset by the patron and if the secessionists' continuation value from fighting exceeds that from the status quo before the continuation value from ceding does, the secessionists will initiate war. Conversely, if the sanctions degrade the secessionists military capabilities sufficiently, it induces the home state to fight.

>>>>

If Condition 1 fails, player $p$ will continue to invest to prevent reunification as in Proposition 1. If Conditions 2 or 3 fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of Condition 3. Note here from Condition 2 that sanctions can induce investment behavior by the patron that was ruled out under the restrictions of Definition 1: the goal of sanctions is to destabilize the Status Quo Equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.

Thus, even if we do not consider the sanctions to impose any direct costs on the home state (though in practice they likely do), it remains ambiguous whether sanctions will benefit the home state. Recall that in the Status Quo Equilibrium, the home state's expected returns from war are lower than from a continuation of the status quo. If sanctions induce the secessionists to play Fight, this is a worse outcome for the home state than if the status quo had been allowed to persist.

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities. The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.

@@

The proof of Proposition 3 is immediate. Although under the conditions of Proposition 1 (i.e. when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting. The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.

Thus, Propositions 2 and 3 imply that sanctions are both wealth destroying and violence increasing. The sanctions destroy wealth directly by damaging the economy of the secessionist region and lowering the secessionists' payoffs from the status quo. If the degradation of status quo payoffs are not offset by the patron and if the secessionists' continuation value from fighting exceeds that from the status quo before the continuation value from ceding does, the secessionists will initiate war. Conversely, if the sanctions degrade the secessionists military capabilities sufficiently, it induces the home state to fight.
